Viscous Transport in Eroding Porous Media

### 2019-10-16 docs/transport/erosion_transport.tex
gave it one last read through

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Integration}
\label{sec:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{Bryan is here}
We use the time stepping method outlined in our previous
work~\citep[see][section 3.3]{qua-moo2018} which we briefly summarize
here.  The erosion rate loses differentiability if the shear stress
changes sign, and this leads to corners developing on $\gamma$ and
numerical instabilities.  Therefore, we modify the erosion rate, $V_\nn$
in equation~\eqref{eqn:erosionModel}, with
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\upi} \kappa - 1 \right),
\end{align}
where $\epsilon \ll 1,$ $\langle \cdot \rangle$ is the spatial average,
$L$ is the length of $\gamma$, and $\kappa$ is the curvature of
$\gamma$.  This new term penalizes regions of high curvature, and the
normalization term guarantees that the regularization only changes the
body's shape, but not its total length.  Finally, to increase the
overall stability of the method, a narrow Gaussian filter is applied to
the erosion rate at each time step.

Rather than tracking the $(x,y)$ coordinates, the {\thL} coordinates are
tracked. In addition, a tangential velocity field is added to each body
so that the discretization points remain equispaced in arclength for all
time.  Time stepping is performed with a second-order Implicit-Explicit
Runge-Kutta method. In particular, the diffusive term corresponding to
the curvature penalization term is discretized implicitly, and all other
terms, which are non-stiff, are treated explicitly.  By using this time
stepping method in conjunction with the Barycentric rule, we stably
evolve the eroding bodies.

To examine the tortuosity and the anomalous dispersion rates
(section~\ref{sec:transport}), we require accurate computation of the
streamlines governed by equation~\eqref{eqn:tracers}. If a low-order
time stepping method is used, or if $\uu(\ss(t))$ is inaccurate, then a
trajectory $\ss(t)$ can unphysically enter a grain, rendering the
trajectory meaningless. However, simply ignoring trajectories that pass
close to a grain could significantly bias the characterization of
transport.  Therefore, we use high-order quadrature and time stepping
methods.  In particular, depending on the proximity of $\ss(t)$ to
$\bd\Omega$ (section~\ref{sec:fmm}), we apply the trapezoid rule or the
Barycentric quadrature rule.  For time stepping, we use a fourth-order
explicit Runge-Kutta method.  By using these high-order methods, we are
able to simulate dynamics very close to the eroding bodies (see
figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}).

Once a collection of trajectories are formed, they are used to quantify
the dispersion and the tortuosity.  We use $N_p = 1000$ streamlines so
that the statistics have converged~\citep{bel-sal-rin1992}.  However, as
described in section~\ref{sec:dispersion}, a reinsertion method is used
to compute trajectories that are sufficiently long to observe an
asymptotic anomalous dispersion rate.  To compute the tortuosity using
the trajectories, we consider trajectories crossing between the two
cross-sections $x=-1$ and $x=1$.  Then, we discretize
equation~\eqref{eqn:tortuosity1} as
\begin{align}
  T = \frac{1}{d}\left(\sum_{i=1}^{N_p} 
    u_1(y_i) \lambda(y_i) \Delta y \right)
  \Bigg/
  \left(\sum_{i=1}^{N_p} u_1(y_i) \Delta y\right), 
%  T = \frac{1}{d}\frac{\displaystyle\sum_{i=1}^{N_p} 
%    u_1(y_i) \lambda(y_i) \Delta y}
%  {\displaystyle\sum_{i=1}^{N_p} u_1(y_i) \Delta y}, 
\end{align}
where $\Delta y = 2/(N_p + 1)$ and $y_i = -1 + i \Delta y$,
$i=1,\ldots,N_p$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Integration}
\label{sec:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We use the time stepping method outlined in our previous
work~\citep[see][section 3.3]{qua-moo2018} which we briefly summarize
here.  The erosion rate loses differentiability if the shear stress
changes sign, and this leads to corners developing on $\gamma$ and
numerical instabilities.  Therefore, we modify the erosion rate, $V_\nn$
in equation~\eqref{eqn:erosionModel}, with
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\upi} \kappa - 1 \right),
\end{align}
where $\epsilon \ll 1,$ $\langle \cdot \rangle$ is the spatial average,
$L$ is the length of $\gamma$, and $\kappa$ is the curvature of
$\gamma$.  The new erosion model penalizes regions of high curvature,
but does not change the total length of each body.  Moreover, to
increase the overall stability of the method, a narrow Gaussian filter
is applied to the erosion rate at each time step.

Rather than tracking the $(x,y)$ coordinates, the {\thL} coordinates are
tracked. In addition, tangential velocity fields are used to maintain
an equispaced discretization. Time stepping is performed with a
second-order Implicit-Explicit Runge-Kutta method. In particular, the
diffusive term corresponding to the curvature penalization term is
discretized implicitly, and all other terms, which are non-stiff, are
treated explicitly.  By using this time stepping method in conjunction
with the Barycentric rule, we stably evolve the eroding bodies.

To examine the tortuosity and the anomalous dispersion rates
(section~\ref{sec:transport}), we require accurate 
streamlines governed by equation~\eqref{eqn:tracers}. If a low-order
time stepping method is used, or if $\uu(\ss(t))$ is inaccurate, then a
trajectory $\ss(t)$ can unphysically enter a grain, rendering the
trajectory meaningless. However, simply ignoring trajectories that pass
close to a grain could significantly bias the characterization of
transport.  Therefore, we use high-order quadrature and time stepping
methods.  In particular, depending on the proximity of $\ss(t)$ to
$\bd\Omega$ (section~\ref{sec:fmm}), we apply the trapezoid rule or the
Barycentric quadrature rule.  For time stepping, we use a fourth-order
explicit Runge-Kutta method.  By using these high-order methods, we are
able to simulate dynamics very close to the eroding bodies (see
figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}).

Once a collection of trajectories are formed, they are used to quantify
the dispersion and the tortuosity.  We use $N_p = 1000$ streamlines so
that the statistics have converged~\citep{bel-sal-rin1992}. As described
in section~\ref{sec:dispersion}, a reinsertion method is used to compute
trajectories that are sufficiently long to observe an asymptotic
anomalous dispersion rate.  To compute the tortuosity using
equation~\eqref{eqn:tortuosity1}, we consider trajectories crossing
between the two cross-sections $x=-1$ and $x=1$, and approximate the
tortuosity with
\begin{align}
  T = \frac{1}{d}\left(\sum_{i=1}^{N_p} 
    u_1(y_i) \lambda(y_i) \Delta y \right)
  \Bigg/
  \left(\sum_{i=1}^{N_p} u_1(y_i) \Delta y\right), 
%  T = \frac{1}{d}\frac{\displaystyle\sum_{i=1}^{N_p} 
%    u_1(y_i) \lambda(y_i) \Delta y}
%  {\displaystyle\sum_{i=1}^{N_p} u_1(y_i) \Delta y}, 
\end{align}
where $\Delta y = 2/(N_p + 1)$ and $y_i = -1 + i \Delta y$,
$i=1,\ldots,N_p$.
